---
title: "Regression-model-final-project"
author: "by fabiobianco"
date: "04/06/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(GGally)
library(datasets)
```


##### Executive Summary

In this paper we are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome) for the `mtcars` dataset. We are particularly interested in the following two questions:

* Is an automatic or manual transmission better for MPG
* Quantify the MPG difference between automatic and manual transmissions


##### Exploratory data analysis
```{r cars}
data("mtcars")
dim(mtcars)
names(mtcars)
str(mtcars)
head(mtcars)
```

Now we display the pairwise relations between variables in the `mtcars` dataset

```{r,  message=FALSE, warning=FALSE}
g = ggpairs(mtcars, lower = list(continuous = wrap("smooth", method = "lm", alpha = 0.3, color = "blue")) )
g

```

In the first columb we can see the relation between the `mpg` variable (outcome) and the other variables (predictors) it seems to be a linear relation (with a positive or negative slope) for every row/predictor and outcome. 

Now we explore the relation between the `mpg` variable (outcome) and the `am` variable (predictor)

```{r}
mtcars1 <- mtcars[mtcars$am == 1,] # mean MPG for manual trasmission system
summary(mtcars1$mpg)
mtcars0 <- mtcars[mtcars$am == 0,] # mean MPG for automatic trasmission system
summary(mtcars0$mpg)
```

```{r}
g = ggplot(mtcars, aes(am, y = mpg, color = am)) + geom_point() + geom_smooth(method = "lm")
g = g + xlab("Trasmission type (0 - Automatic, 1 - Manual)") + ylab("Miles per gallon (MPG)")
g = g + labs(title = paste("MPG vs Trasmission type"))
g
```


##### Regression Models

Now fit a multivariable linear regression model for the `mtcars` dataset

```{r, Multivariable linear regression}
fitall <- lm(mpg ~ . , data = mtcars)
summary(fitall)
```

The  `fitall` model accounts for 81% of the variance as noted by the adjusted Rsquared value.

Now we investigate the relationship between trasmission type `am`(predictor) and miles per gallon `MPG` (outcome)

```{r, linear regression analysis}
fit <- lm(mpg ~ am,  data = mtcars)
summary(fit)
```

The `fit` model accounts for 34% of the variance as noted by the adjusted Rsquared value. There seems other predictors have some impact on MPG. The `fit` model predicts an extra 7.245 mpg consuption for manual trasmission veichle versus automatic trasmission veichle. Examining the regression output value, we can see that the p-value for `am` is very clode to zero, indicating there is strong evidence that the coefficient is different fro zero when using this one-variable model.


As a final step we search the model that best fit the data. 

```{r, include=FALSE}
bestmodel <- step(fitall)
```

```{r}
summary(bestmodel)
```

The `bestmodel` accounts for 83% of the variance as noted by the adjusted Rsquared value. 


##### Appendix

Residual for **Full model** (all predictors)
```{r}
par(mfrow=c(2, 2))
plot(fitall)
```

Residual for **single-variable model** (only `am` predictor)

```{r}
par(mfrow=c(2,2))
plot(fit)
```

Residual for **best model** (only `am`, `wt`, `qsec` predictors)

```{r}
par(mfrow=c(2, 2))
plot(bestmodel)
```

